Efficient Coalition Formation \\ for Web Services

### 2013-12-02 papers/coalition-journal-extension/coalition_journal_bak2.tex
fixes-2

\subsection{Cooperative Game Concepts}
Cooperative game is a branch of game theory that studies
strategies of self-interested entities or agents in a setting
where those agents can increase their payoff by binding agreements
and cooperating in groups. We let $N$ be a set of players which can form a group called a $coalition$. A
\emph{coalitional game} is a pair $G = (N, v)$, where $v$ is called
a \emph{characteristic function} $v: 2^N \to \mathbb{R}$, mapping the set of players of the
coalition to a real number $v(N)$, the worth of $N$. This number
usually represents the output or payoff or again the performance
of these players working together as coalition.  If a coalition
$S$ is formed, then it can divide its worth, $v(N)$ in any
possible way among its members. The payoff vector $x \in
\mathbb{R}^N$ is the amount of payoff being distributed among the
members of the coalition $N$. The payoff vector satisfies two
conditions:

\begin{itemize}
    \item $x_i \geq 0$ for all $i \in N$, and
    \item $\sum_{i \in N} x_i \leq v(N)$
\end{itemize}

The second criteria is called the \emph{feasibility} condition,
according to which, the payoff for each agent cannot be more than
the coalition total gain. A payoff vector is also \emph{efficient}
if the payoff obtained by a coalition is distributed amongst the
coalition members: $\sum_{i \in N} x_i = v(N)$. This definition of
the characteristic function works in \emph{transferable utility}
(TU) settings, where utility (i.e., payoff) is transferable from
one player to another, or in other words, players have common
currency and a unit of income that is worth the same for all players
\cite{myerson1991game}.

@@

\emph{Shapley value} is a unique and fair solution concept for
payoff distribution among the members of the coalition. It
basically rewards members with the amount of average marginal contribution
they add to the coalition. It checks the contribution of member $i$ by adding the agent, to all possible subsets 
of coalitions $S$, where ${S \subseteq N\left\backslash\left\{i\right\}$. If he is added to the set $S$, his 
contribution to the coalition is $v(S \cup \left\{i\right\}) - v(S)$. Average marginal contribution of agent $i$'s 
is calculated by averaging this value over all possible subsets of $N$, in $Shapley value$ equation (\ref{eq:shapley}).

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
% \caption{Three web services}
\label{example_3_2}
\centering
\begin{tabular}{c c || c c}
\hline
Community & Worth & Community & Worth\\
\hline
$\left\{C_{ms_1}\right\}$ & 0 & $\left\{C_{ms_2}\right\}$ & 0\\
$\left\{C_{ms_1}, ws_1\right\}$ & 8 & $\left\{C_{ms_2}, ws_1\right\}$ & 8\\
$\left\{C_{ms_1}, ws_2\right\}$ & 16 & $\left\{C_{ms_2}, ws_2\right\}$ & 16\\
$\left\{C_{ms_1}, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_3\right\}$ & 24\\
$\left\{C_{ms_1}, ws_1, ws_2\right\}$ & 16 & $\left\{C_{ms_2}, ws_1, ws_2\right\}$ & 24\\
$\left\{C_{ms_1}, ws_1, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_1, ws_3\right\}$ & 32\\
$\left\{C_{ms_1}, ws_2, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_2, ws_3\right\}$ & 32\\
$\left\{C_{ms_1}, ws_1, ws_2, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_1, ws_2, ws_3\right\}$ & 32\\
$\left\{C_{ms_1}, C_{ms_2}, ...\right\}$ & 0 & $\left\{ws_1\right\}$ & 6.8\\
$\left\{ws_2\right\}$ & 4.2 & $\left\{ws_3\right\}$ & 6.8\\
\hline
Community $R_{C_1}$: 20 \\ Community $R_{C_2}$: 40\\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In Example 3, we consider the case of having different communities with different market share, ${R_C}$ values. Web services also have a small share of market independently, providing them with a small task pull. In these kind of scenarios, the solution considers individual maximization of payoff and also the total worth of all communities which represents the \emph{social welfare}. In this example the most efficient partition of web services is earned by having two coalitions of $\left\{C_{master_1}, ws_2\right\}$ and $\left\{C_{master_2}, ws_1, ws_3\right\}$, which yields a total value of $32 + 16 = 48$. In these types of scenarios, the goal is to reach stability, adopting a distributed approach where all players have the power of choice on the decision of whether or not they join a coalition. The communities usually start the game having some established members, encountering new web services, the communities may exchange web services and new web services would join them having at least one player gaining utility, without hurting any other participant. In this example if we initially having two coalitions of $\left\{C_{master_1}, ws_2\right\}$ and $\left\{C_{master_2}, ws_1\right\}$ and a ${ws_3}$ as new web service, ${ws_3}$ joining ${C_{master_1}}$ would hurt at least itself or $ws_2$, however ${ws_3}$ joining ${C_{master_2}}$ would not hurt any participants and ${ws_3}$ would earn more within the community and the community will have enough web services performing the incoming tasks from users.

@@

As mentioned in previous section, the worth and output of a
community  is a function of the
throughput and provided QoS of its web service members. If the throughput rate is more than
the master's input request rate, it means the web services inside
the community are capable of serving more requests than the
demand. Considering this factor, the valuation function is
designed to balance the output performance so that it matches the
exact throughput rate and QoS the web service can provide within
the particular community.
%** In the case where the limited tasks are distributed among web services uniformly, the value of coalition would be the proportion of the average QoS times their throughput to rate of available requests. **

@@

Considering $S \subseteq T$:
\setlength{\abovedisplayshortskip}{2pt}
\begin{gather*}
T \cup \left\{i\right\} = (S \cup \left\{i\right\}) \cup T
\\
S = (S \cup \left\{i\right\}) \cap T
\end{gather*}

@@

Our algorithm works as follows. Given an  established community
with a master and some member web services, a web service would send a \emph{join request} to join  the
community. Ideally, the \emph{core}
or \emph{$\epsilon$-core} stability of the group having this new
member should be analyzed. As the normal core membership algorithm is computationally
intractable, we exploit Proposition \ref{proposition1} and Equation
\ref{eq:convex_snow} to check the convexity of our game having
characteristic function where the new member is added. In the
equation, let $C$  be our community members before
having the new web service join the community. Let  ${i}$ be the new web
service, and then verify the equation for $S$, setting $ S = T /
W1 $ where $W1$ is the set of all possible subsets of the set $N$
having the size $1$. We can relax the equation a bit by adding a
constant $\epsilon$ to the left side of the equation. We call this
method \emph{Depth-1 Convex-Checker} algorithm. If the equation is
satisfied for all $W1$, we let the new web service join our
community, since the web service will contribute positively enough
to make our new community stable. Since only subsets of size $1$
are checked, the following Proposition holds.

@@

It is possible to develop  algorithm that continues the
verification of this condition against subsets of size $3$,
$4$, etc. until the algorithm gets interrupted.

@@

In this scenario, we consider multiple communities managed by
multiple master web services, each of which is providing
independent request pools (see Figure \ref{fig_sim2}). Identical
to the first scenario, master web services form coalitions with
web services. We use coalition structure formation methods to
partition web services into non-empty disjoint coalition
structures. As mentioned in Section \ref{sec:coalition}, the used
algorithms in \cite{Sandholm1999209,
DBLP:conf/ijcai/GrecoMPS11,DBLP:conf/ijcai/RahwanMJ11} try to
solve key fundamental problems of what coalitions to form, and how
to divide the payoffs among the collaborators.

@@

In this scenario, as new web services are discovered and get ready
to join communities, our algorithm keeps merging and splitting
partitions based on the preference function. The decision to merge
or split is based on the fact that all players must benefit. The
new web services will merge with communities if $all$ the players
are able to improve their individual payoff, and some web services
may split from old communities, if splitting does not decrease the
payoff of \emph{any} web service of the community. According to
\cite{DBLP:journals/corr/abs-cs-0605132}, this sequence of merging
and splitting will converge to a final partition, where web
services cannot improve their payoff. More details of this
algorithm and analysis of generic solutions on coalition formation
games are described in \cite{DBLP:journals/igtr/AptW09}.

\subsection{Taxation, Subsiding and Community Stability}\label{s:tax}

We discussed $core$ as one of the prominent solution concepts in cooperative games. Working together, completing tasks and generating revenue, agents need to distribute the gain in a way no agents would gain more by having their own group. However, in most cases the core of a game is empty so we introduced the \emph{$\epsilon$-core} concept, where agents would only earn a minimal amount of $\epsilon$ by deviating. Stability is an attractive property for communities. In addition, communities would benefit by having slightly more web services then the exact number of web services satisfying task rate cap. This is because  there is always a possibility that the agents may leave the community or they may under perform and degrade the quality values they were initially performing with.

One way for communities to ensure  stability could be  by applying a tax, $\epsilon$ amount of cost for web services changing communities which would make deviation a costly act. However this would require all community coordinators to agree on a same amount of taxation, being governed by some external entities or web services would join communities with lesser amount of tax. Another viable solution to our scenario is to stabilize the game using external subsidies. The reason a game is not stable is that the community is not making enough revenue to allocate  enough gain to the players. A community master can subside its community with a constant coefficient value of $\lambda$.

Rewarding a community with high number of quality participants with $\lambda v(C)$, where $\lambda \leq 1$.

Obviously with a big number of $\lambda$, it is always possible to stabilize the community. However, this can be a costly act for community masters so they are interested in the minimum subside value of $\lambda$ making the game stable. Subsiding or taxing in order to reduce the bargaining power of sub coalitions are called $taxation$ \cite{eps346856} methods.

@@

One of the key criteria reflecting the performance of web service
coalitions is the user satisfaction. User satisfaction can be
measured in terms of quality and quantity of requests (or tasks)
successfully answered by the communities. We initiated the
communities with few web services, then let rejecting and
accepting random web services go for a short number of iterations.
After that, we started the request distribution for the communities
and let them allocate requests among member web services.
Thereafter, we measured the average output performance of tasks in
communities following different methods.

@@

For the \emph{optimal core} method we have used the well known \emph{$\epsilon$-core} method as the
taxtation method to relax the core condition to help communities, attract web services.
We have assigned $\epsilon$ to 15\% of total community worth, $\epsilon = 0.15 \times v(C)$, which allows
subsets of the coalition to gain maximum 15\% of $v(C)$. In the
\emph{optimal $\epsilon$-core} method, we capped the coalition
size to 25 web services, since the method is computationally
intractable as number of web services increase and anything more than that would make it
impractical to run in our simulations. In the other methods, there
were no cap on size of the community and we had communities of
size 60 web services at some points. In this scenario our community receives 30 tasks on average per iteration, from users. The community, after the task distribution process on each iteration, will reevaluate QoS metrics of its members and can check for new membership requests. Web services may join or leave the community between iterations. The results show that our
\emph{depth-2 convex checker} method is performing better compared
to the other methods and its performance is close to optimal
\emph{$\epsilon$-core} method. Our \emph{depth-1 convex checker}
and the \emph{3-Way Satisfaction} method, are also performing well.

@@

In Figure \ref{performancemany}, we compare our \emph{Web Services
and many Communities} scenario with a method which ignores QoS
parameters and forms coalitions by allowing web services to join
only if they have enough requests for themselves. In other words,
web services can join a community when the request rate is less
than the throughput of all the member web services. We name this
method \emph{Random Formation} and use it as a benchmark for our
QoS-aware coalition formation process. In this scenario, each user individually generates randomly 
between 0 to 10 number of tasks per iteration, then the users target a community and direct their requests to the chosen community. As the results illustrate,
our method forms better coalitions of web services improving
performance and satisfaction for both web services and coalitions.

>>>>

\subsection{Cooperative Game Concepts}
Cooperative game is a branch of game theory that studies
strategies of self-interested entities or agents in a setting
where those agents can increase their payoff by binding agreements
and cooperating in groups. We let $N$ be a set of players. Any
subset $S$ of $N$ can form a group called $coalition$. A
\emph{coalitional game} is a pair $G = (N, v)$, where $v$ is called
a \emph{characteristic function} $v: 2^N \to \mathbb{R}$, mapping the set of players of the
coalition to a real number $v(S)$, the worth of $S$. This number
usually represents the output or payoff or again the performance
of these players working together as coalition.  If a coalition
$S$ is formed, then it can divide its worth, $v(S)$ in any
possible way among its members. The payoff vector $x \in
\mathbb{R}^S$ is the amount of payoff being distributed among the
members of the coalition $S$. The payoff vector satisfies two
conditions:

\begin{itemize}
    \item $x_i \geq 0$ for all $i \in N$, and
    \item $\sum_{i \in S} x_i \leq v(S)$
\end{itemize}

The second criteria is called the \emph{feasibility} condition,
according to which, the payoff for each agent cannot be more than
the coalition total gain. A payoff vector is also \emph{efficient}
if the payoff obtained by a coalition is distributed amongst the
coalition members: $\sum_{i \in S} x_i = v(S)$. This definition of
the characteristic function works in \emph{transferable utility}
(TU) settings, where utility (i.e., payoff) is transferable from
one player to another, or in other words, players have common
currency and a unit of income that is worth the same for all players
\cite{myerson1991game}.

@@

\emph{Shapley value} is a unique and fair solution concept for
payoff distribution among the members of the coalition. It
basically rewards members with the amount of marginal contribution
they have to the coalition.

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
% \caption{Three web services}
\label{example_3_2}
\centering
\begin{tabular}{c c || c c}
\hline
Community & Worth & Community & Worth\\
\hline
$\left\{C_{ms}_1\right\}$ & 0 & $\left\{C_{ms}_2\right\}$ & 0\\
$\left\{C_{ms}_1, ws_1\right\}$ & 8 & $\left\{C_{ms}_2, ws_1\right\}$ & 8\\
$\left\{C_{ms}_1, ws_2\right\}$ & 16 & $\left\{C_{ms}_2, ws_2\right\}$ & 16\\
$\left\{C_{ms}_1, ws_3\right\}$ & 16 & $\left\{C_{ms}_2, ws_3\right\}$ & 24\\
$\left\{C_{ms}_1, ws_1, ws_2\right\}$ & 16 & $\left\{C_{ms}_2, ws_1, ws_2\right\}$ & 24\\
$\left\{C_{ms}_1, ws_1, ws_3\right\}$ & 16 & $\left\{C_{ms}_2, ws_1, ws_3\right\}$ & 32\\
$\left\{C_{ms}_1, ws_2, ws_3\right\}$ & 16 & $\left\{C_{ms}_2, ws_2, ws_3\right\}$ & 32\\
$\left\{C_{ms}_1, ws_1, ws_2, ws_3\right\}$ & 16 & $\left\{C_{ms}_2, ws_1, ws_2, ws_3\right\}$ & 32\\
$\left\{C_{ms}_1, C_{ms}_2, ...\right\}$ & 0 & $\left\{ws_1\right\}$ & 6.8\\
$\left\{ws_2\right\}$ & 4.2 & $\left\{ws_3\right\}$ & 6.8\\
\hline
Community $R_{C_1}$: 20 \\ Community $R_{C_2}$: 40\\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In Example 3, we consider the case of having different communities with different market share, ${R_C}$ values. Web services also have a small share of market independently, providing them with a small task pull. In these kind of scenarios, the solution considers individual maximization of payoff and also the total worth of all communities which represents the \emph{social welfare}. In this example the most efficient partition of web services is earned by having two coalitions of $\left\{C_{master}_1, ws_2\right\}$ and $\left\{C_{master}_2, ws_1, ws_3\right\}$, which yields a total value of $32 + 16 = 48$. In these types of scenarios, the goal is to reach stability, adopting a distributed approach where all players have the power of choice on the decision of whether or not they join a coalition. The communities usually start the game having some established members, encountering new web services, the communities may exchange web services and new web services would join them having at least one player gaining utility, without hurting any other participant. In this example if we initially having two coalitions of $\left\{C_{master}_1, ws_2\right\}$ and $\left\{C_{master}_2, ws_1\right\}$ and a ${ws_3}$ as new web service, ${ws_3}$ joining ${C_{master}_1}$ would hurt at least itself or $ws_2$, however ${ws_3}$ joining ${C_{master}_2}$ would not hurt any participants and ${ws_3}$ would earn more within the community and the community will have enough web services performing the incoming tasks from users.

@@

As mentioned in previous section, the worth and output of a
community of web services is a function of the
throughput and provided QoS. If the throughput rate is more than
the master's input request rate, it means the web services inside
the community are capable of serving more requests than the
demand. Considering this factor, the valuation function is
designed to balance the output performance so that it matches the
exact throughput rate and QoS the web service can provide within
the particular community.
%** In the case where the limited tasks are distributed among web services uniformly, the value of coalition would be the proportion of the average QoS times their throughput to rate of available requests. **

@@

Considering $S \subseteq T$:
\setlength{\abovedisplayshortskip}{2pt}
\begin{gather*}
S \cup \left\{i\right\} = (S \cup \left\{i\right\}) \cup T
\\
T = (S \cup \left\{i\right\}) \cap T
\end{gather*}

@@

Our algorithm works as follows. We have an established community
with a master and some member web services already residing in the
community. A web service would send a join request to the
community. The ideal solution would be analyzing the \emph{core}
or \emph{$\epsilon$-core} stability of the group having this new
member. As the normal core membership algorithm is computationally
intractable, we exploit Proposition \ref{proposition1} and Equation
\ref{eq:convex_snow} to check the convexity of our game having
characteristic function where the new member is added. In the
equation, we set $C$ to be our community members $(C)$ before
having the new web service. We assign ${i}$ to the new web
service, and then verify the equation for $S$, setting $ S = T /
W1 $ where $W1$ is the set of all possible subsets of the set $N$
having the size $1$. We can relax the equation a bit by adding a
constant $\epsilon$ to the left side of the equation. We call this
method \emph{Depth-1 Convex-Checker} algorithm. If the equation is
satisfied for all $W1$, we let the new web service join our
community, since the web service will contribute positively enough
to make our new community stable. Since only subsets of size $1$
are checked, the following Proposition holds.

@@

It is possible to develop an anytime algorithm by continuing the
verification of this condition against all subsets of size $3$,
$4$, etc. until the algorithm gets interrupted.

@@

In this scenario, we consider multiple communities managed by
multiple master web services, each of which is providing
independent request pools (see Figure \ref{fig_sim2}). Identical
to the first scenario, master web services form coalitions with
web services. We use coalition structure formation methods to
partition web services into non-empty disjoint coalition
structures. As mentioned in Section \ref{sec:coalition}, the used
algorithms \cite{Sandholm1999209,
DBLP:conf/ijcai/GrecoMPS11,DBLP:conf/ijcai/RahwanMJ11} try to
solve key fundamental problems of what coalitions to form, and how
to divide the payoffs among the collaborators.

@@

In this scenario, as new web services are discovered and get ready
to join communities, our algorithm keeps merging and splitting
partitions based on the preference function. The decision to merge
or split is based on the fact that all players must benefit. The
new web services will merge with communities if $all$ the players
are able to improve their individual payoff, and some web services
may split from old communities, if splitting does not decrease the
payoff of any web service of the community. According to
\cite{DBLP:journals/corr/abs-cs-0605132}, this sequence of merging
and splitting will converge to a final partition, where web
services cannot improve their payoff. More details of this
algorithm and analysis of generic solutions on coalition formation
games are described in \cite{DBLP:journals/igtr/AptW09}.

\subsection{Taxtation, Subsiding and Community Stability}\label{s:tax}

We discuessed $core$ as one of the preminent solution concepts in cooperative games. Working together, completing tasks and generating revenue, agents need to distribute the gain in a way no agents would gain more by having their own group. However, in most cases the core of a game is empty so we introduced the \emph{$\epsilon$-core} concept, where agents would only earn a minimal ammount of $\epsilon$ by deviating. Comuunity stability is an attractice property for communities, also communities would benefit by having slightly more web services then having the exact number of web services satisfying task rate cap. One reason is that there is always the possibility that the agents may leave the community or they may under perform from the quality values they were initially performing with.

One way for communities to ensure community stability could be applying a tax, $\epsilon$ ammount of cost for web services changing communities which would make deviation a costly act. However this would require all community coordinators to agree on same ammount of taxtation, being governed by some external entities or web services would join communities with lesser ammount of tax. Another viable solution in our scenario is to stabilize the game using external subsidies. The reason a game is not stable is that the community is not making enough revenue to allocate the players with enough gain. A community master can subside its community with a constant coefficient value of $\lambda$. Rewarding a community with high number of quality participants with $\lambda v(C)$, where $\lambda \leq 1$. Obviously with a big number of $\lambda$ it is always possible to stabilize the community. However, this can be a costly act for community masters so they are interested in the minimum subside value of $\lambda$ making the game stable. Subsiding or taxting in order to reduce the bargaining power of sub coalitions are called $taxtation$ \cite{eps346856} methods.

@@

One of the key criteria reflecting the performance of web service
coalitions is the user satisfaction. User satisfaction can be
measured in terms of quality and quantity of requests (or tasks)
successfully answered by the communities. We initiate the
communities with few web services, then let rejecting and
accepting random web services go for a short number of iterations.
After that, we start the request distribution for the communities
and let them allocate requests among member web services.
Thereafter, we measure the average output performance of tasks in
communities following different methods.

@@

For the \emph{optimal core} method we have used the well known \emph{$\epsilon$-core} method as the
taxtation method to relax the core condition to help communities, attract web services.
We have assigned $\epsilon$ to 15\% of total community worth, $\epsilon = 0.15 \times v(C)$, which allows
subsets of the coalition to gain maximum 15\% of $v(C)$. In the
\emph{optimal $\epsilon$-core} method, we capped the coalition
size to 25 web services, since the method is computationally
intractable as number of web services increase and anything more than that would make it
impractical to run in our simulations. In the other methods, there
were no cap on size of the community and we had communities of
size 60 web services at some points. The results show that our
\emph{depth-2 convex checker} method is performing better compared
to the other methods and its performance is close to optimal
\emph{$\epsilon$-core} method. Our \emph{depth-1 convex checker}
and the \emph{3-Way Satisfaction} method, are also performing well.

@@

In Figure \ref{performancemany}, we compare our \emph{Web Services
and many Communities} scenario with a method which ignores QoS
parameters and forms coalitions by allowing web services to join
only if they have enough requests for themselves. In other words,
web services can join a community when the request rate is less
than the throughput of all the member web services. We name this
method \emph{Random Formation} and use it as a benchmark for our
QoS-aware coalition formation process. As the results illustrate,
our method forms better coalitions of web services improving
performance and satisfaction for both web services and coalitions.
